---
title: "WUP_04"
author: "Colton Beardsley"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using ChatGPT and CandP helper files to compare numerical data and categorical data in the GSS data set. we will compare WORDSUM and how it was effected by education level.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

# Create an elegant scatterplot

```{r include=TRUE,echo=FALSE}
ggplot(data = mydata, aes(x = DEGREE, y = WORDSUM) ) +
geom_boxplot ( notch=FALSE , outlier.colour = "red" , outlier.size = 2 , fill = "darkgray" ) +
stat_boxplot ( geom = 'errorbar' ) +
labs ( x = "DEGREE" , y = "WORDSUM SCORE" , title = "WORDSUM EXPLAINED BY WORDSUM" ) +
theme ( plot.title = element_text (hjust = 0.5, size = 16) )
```

From this box plot we can tell that the entire IQR of the High School and LT High School is underneath the median of all of the college groups.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```

The mean and median of the LT High school and the high school subcategories are for lower than means and medians of the university level data points.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
summary(model1)
```

```{r include=TRUE,echo=FALSE}
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1,conf.level = 0.99),las=2)
```

The results from both the TukeyHSD and aov test shows that there are significant differences in the different DEGREE categories. This can be best seen when we apply a high confidence level to the results, as we did with the TukeyHSD model that had a 99% confidence level, which showd a significant difference in the HIGH SCHOOL-GRADUATE and LT HIGH SCHOOL-GRADUATE. The low p-value from the avo test is a very small value, 0.000236. This means that there is a statistically significant results, which is evidence against the null hypothesis.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

From the first box plot we could tell that the entire IQR of the High School and LT High School was underneath the median of all of the college groups. By itself this is not enough to prove that there was a significant statistically difference, but it did serve as a good starting point of analysis. From this point we went on to look at the favstats test in order to further compare the data between the various categories. This proved to show that the mean and median the LT High School and the High School subcategories are for lower than means and medians of the university level data points. We then ran an aov test, which showed that there was a very small p-value of 0.000236, showing that there was a statically significant result in the data. This is further proved by running the TukeyHSD test, which showed a statically difference with a 99% confidence level applied to it.

This all proves to show that there prove the $H_a$ and that there is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE, that being the difference between the average Highschool WORDSUM test scores and the average College WORDSUM test scores.

